MSstatsPTM: Statistical relative quantification of post-translational modifications in bottom-up mass spectrometry-based proteomics

### 2022-07-30 main/MSstatsPTM_Main.tex
More evaluation

\figref{sim1_fdr}(a) summarizes the FDR \todo{I am not sure we should call this FDR. By definition FDR is a population-level quantity (it is an expected value of the proportion of false positives. Here it is an observed value). Maybe replace with 1-PPV?} reported on Computer Simulation 1 dataset by the unadjusted and adjusted versions of methods \todo{XXX}. All the analyses were performed to control the FDR at at most 5\%.  The  simulation mimics a ``clean" label-free experiment, not compromised by issues such as deviations from model assumptions, missing values and outliers. Yet, even in these favorable situations, all the versions of the models that did not adjust for confounding from changes in overall protein abundance produced an excessive number of false positives. The versions of the models that accounted for this confounding produced much better calibrated error rates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{In noisy simulations, MSstatsPTM better calibrated the true fold change over the existing methods}

@@

\begin{figure}[ht]
\centering
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_fc.png}
\caption{}
\label{fig:spikein_boxplot}
\end{subfigure}
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_volcano.png}
\caption{}
\label{fig:spikein_prop_volcano}
\end{subfigure}
\caption{Dataset 3: Spike-in benchmark - Ubiquitination - Label-free. a) Before adjustment the fold change of the spike-in peptides' were systematically different from the expected fold change in all models. After adjustment, this systemic difference was removed, however the inner quartile range of the Limma and ANOVA models was wider than the proposed method. b) Before adjustment the spike-in peptides (colored red) did not follow the expected log fold change; after adjustment, the spike-in peptides were more in line with expectation. Using Limma, the spike-in peptides followed the expected log fold change after adjustment, however the majority of spike-in peptides did not have a significant adjusted p-value.}
\label{fig:spikein_volcano}
\end{figure}

>>>>

\figref{sim1_fdr}(a) summarizes the FDR \todo{I am not sure we should call this FDR. By definition FDR is a population-level quantity (it is an expected value of the proportion of false positives. Here it is an observed value). Maybe replace with 1-PPV?} reported on Computer Simulation 1 dataset by the unadjusted and adjusted versions of methods \todo{XXX}. All the analyses were performed to control the FDR at at most 5\%.  The  simulation mimics a ``clean" label-free experiment, not compromised by issues such as deviations from model assumptions, missing values and outliers. Yet, even under these favorable circumstances, the models that did not adjust for confounding from changes in overall protein abundance produced an excessive number of false positives. The versions of the models that accounted for the confounding produced error rates that were much better calibrated at the desired level. Since the simulation mimicked the assumptions of \todo{model XXX}, the error rates of this model were calibrated best.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{In noisy simulations, MSstatsPTM improved the estimation of fold change over the existing methods \todo{Hmm... Fig 4b is not about the fold change... Rephrase?}}

\todo{I suggest this order for this section. (Fig 4b has multiple panels, but the current narrative doesn't make it quite clear what is the difference between the panels) Fig 4b summarizes the results on Sim 2 dataset. The dataset is more realistic, and includes XXX. For low noise, methods compared in terms of sample size and number of conditions XXX. As the noise increased, XXX. The difference is due primarily to the robust nature of summarization in MSstats. Supplementary Fig XXX details the estimation of fold change}

@@

\begin{figure}[ht]
\centering
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_fc.png}
\caption{}
\label{fig:spikein_boxplot}
\end{subfigure}
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_volcano.png}
\caption{}
\label{fig:spikein_prop_volcano}
\end{subfigure}
\caption{Dataset 3: Spike-in benchmark - Ubiquitination - Label-free. a) Before adjustment the fold change of the spike-in peptides' were systematically different from the expected fold change in all models. After adjustment, this systemic difference was removed, however the inner quartile range of the Limma and ANOVA models was wider than the proposed method. b) \todo{The title of the figure reads as if it is accuracy-adjusted. Maybe rephrase?} Before adjustment the spike-in peptides (colored red) did not follow the expected log fold change; after adjustment, the spike-in peptides were more in line with expectation. Using Limma, the spike-in peptides followed the expected log fold change after adjustment, however the majority of spike-in peptides did not have a significant adjusted p-value.}
\label{fig:spikein_volcano}
\end{figure}
